---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# music

<!-- badges: start -->
<!-- badges: end -->

The goal of `music` is double:

- explore music theory through opinionated declarative functions;
- explore R programming and demonstrate some development skills.

See the [Extra](#extra) section for a more detailed description.


## Installation

You can install the development version of music like so:

``` r
#install.packages("pak")
pak::pkg_install("c-marc/music")
```


## Example

1. Start with notes C, F and A, and assume root note is D.
2. Add the role each of these note play.
3. Transform the index so that it now relates to half-tones.
4. Show the first _inversion_ of that chord (starting on its second note).
5. Compute intervals.

```{r example, echo = FALSE}
library(music)

start_with_some_notes(c("C", "F", "A"), root_note = "D") |>
  add_roles() |>
  as_12_TET() |>
  set_first(rank = 2) |>
  dplyr::mutate(intervals = c(0, diff(idx)))
```
This is (D)-7/A: rootless D minor 7 played over A...


## Extra {#extra}

### Music theory

This is an interesting subject to play with for the following reasons:

- It involves a meaningful subject to play with, and one can actually learn some things while playing with data and functions. Yet no specific knowledge is required here, and the implementation is simple enough to be used in an abstract way.

- In the scope of this package, it involves two supplementary interesting features for exploration learning and/or teaching programming skills:

  1. It relies on **simple blocks**. Each of these blocks is quite small (generally up to 12 rows and 1-3 columns). Each of these block can be used on its own or combined with the others.
  
  2. It involves **strong logic**. That means there's a lot to experiment with, from transforming to joining these blocks. Or rebuilding some just by using the others.
  

### Programming

#### General goals

This package has the following general goals:

- to provide a framework with declarative and pure functions that can be used intuitively with basic *tidyverse* skills 
- to provide a carefully crafted package:
  - strong checking of users inputs, and informative messaging;
  - decent documentation;
  - nice and organised coding.


#### Programming topics or skills

It implements:

- OOP in S3 framework, with subclasses of `vctrs` and `tibbles`;
- pretty interactions with `cli`;
- some features requiring programming with _tidy-selection_, _data-masking_, and _glue injections_;
- recursive functions;
- dangerous arithmetic (especially for non interactive use);
- unit testing, and safe programming by being quite strict with types, objects and arguments checking...


#### Design choices

Main design choices are:


##### For objects:

  - implement a powerful index class that can deal with periodicity and be transformed;
  - implement a subclass of a tibble with that index as a column;
  
  
##### For functions

We define 3 levels:

  1. Low level functions are internal...
  2. Mild level functions are in the NAMESPACE: they make available some operations to the user. They require some knowledge to be called meaningfully. And they usually must be sequenced with each other, with joining, and arranging... They require to explicitly unlock the data.
  3. Higher-order functions should work without much thinking from the user and show strong stability over time, hiding potential underlying changes. They can operate on locked data.
  
  
##### For data

We experiment an idea of data _as functions_. This is very simple in that it just wraps data objects behind a proxy function. Our motivations for such a choice are that:

  - this _shielding_ of data behind function brings safety for interactive use. Even though our functions are pure, this encourages the user to call reproducible data (blocks). It's just an assignment away. But as most problems should be solvable in just a few lines, these original blocks do not need to be tempered with. An intermediary assignment will sometimes makes a solution clearer, but most piping sequence are encouraged to start with these blocks.
  - this blurs the distinction between hard coded data and some blocks that can be built on the fly. Something we want to take advantage of. This allows extensions without actually adding data: other _blocks_ can be proposed as arguments through limited getting functions; these can be a view of a block, pre-assembled blocks, modified blocks, etc... This should fit with our approach of using a limited set of clear and intuitive verbs. This functional approach wraps the OOP parts (it actually promotes the top level methods of the richest objects).
