---
title: "str_fun"
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(music)
library(stringr)
```


### With string manipulation

Major penta is diatonic minus 4th and 7th:

```{r}
c("CDEFGAB") %>% str_replace_all("[FB]","")
```

Remove 3rd and 7th, duplicate and pick 5 notes starting on F:

```{r}
c("CDEFGAB") %>% 
  str_replace_all("[EB]","") %>%
  str_dup(2) %>%
  str_extract("F.{4}") 
```
Same idea:

```{r}
c("CDEFGAB") %>% 
  str_replace_all("[CF]","") %>%
  str_dup(2) %>%
  str_extract("G.{4}")
```

... but dealing with # and b is harder...


```{r}
x <-  letters[1:7] %>% str_flatten()
str_replace(x,"(.{5})(.*)","\\2\\1")
```
```{r}
str_replace(x,"(.{2})(.*)","\\2\\1")
```

```{r}
str_replace("CDEFGAB","(.)(.)(.)(.)(.)(.)(.)", "\\1\\5\\2")
```

```{r}
str_replace(
  "CC#DD#EFF#GG#AA#B",
  str_dup("(.[#b]?)",12), "\\1\\8\\3")
```

```{r}
wild <- str_c("\\",((0:11)*7)%%12+1) %>% str_flatten()
wild
```

```{r}
str_replace(
  "CC#DD#EFF#GG#AA#B",
  str_dup("(.[#b]?)",12), wild)
```

An idea could be to use a trick, for example lowercase for bemols:

And `_` for missing: 

- 12 `CdDeEFgGaAbB`
- Cm triad:`C---e----G----`
- C Maj: `C-D-EF-G-A-B`
- D dorian: `D-EF-G-A-BC-`
- Circle of fifths: `CGDEABgdeabF`
- C Maj: `CGDEAB-----F`
- Penta C Maj: `CGDEA------`
- Penta G Blues Maj: `GDEA------C`
