---
title: "modes_7"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modes_7}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(music)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
```


## Explore all 7 modes at once

Add roles to each mode:

```{r}
demo <- show_mode(mode_size = 7) |>
  rowwise() |>
  mutate(data = start_with_mode(mode_size = 7, degree = degree) |>
           add_roles() |>
           as_12_TET() |>
           list())
```

Pivot wider with roles:

```{r}
out <- demo |>
  unnest(data) |>
  mutate(idx = as.integer(idx)) |>
  pivot_wider(names_from = idx, values_from = role) |>
  select("degree", "name", as.character(0:11))

out
```

Or order modes by building order:

```{r}
tibble(degree = c(4, 1, 5, 2, 6, 3, 7)) |>
  left_join(out, by = "degree")
```

To show how this simply aligns in some appropriate coordinates, we just catch results before transforming in 12-TET:

```{r}
demo2 <- get_mode(mode_size = 7) |>
  rename(rank = idx) |>
  rowwise() |>
  mutate(data = start_with_mode(mode_size = 7, degree = degree) |> 
              add_roles() |>
              list())
```

```{r}
out <- demo2 |>
  unnest(data) |>
  mutate(idx = as.integer(idx)) |>
  pivot_wider(names_from = idx, values_from = role) |>
  select("degree", "name", paste0(-6:6))

out
```


## Other ways to construct or understand modes

### Direct slicing

We start with values of an index:

- sequence of 7 consecutive values on the circle of fifths.
- starting on -1 for the ionian mode
- here we directly set the index to a 12-TET
- the important parameter is to set the unit so that it represents fifths for the specified period (`7`; on can use the helper `fifth(period = 12L)`)


```{r}
ionian <- -1
start_with_some_values((0:6) + ionian, period = 12L, unit = 7L) |>
  add_roles()
```

### Sequential building

Here we demonstrate how 7-modes can be seen as cumulative alterations.

```{r}
# Start on lydian mode (fully augmented)
lydian <- start_with_mode(degree = 4) |> add_roles()

# We need to sequentially diminish roles
xx <- lydian |>
  filter(role != "P1") |> 
  pull("role") |> 
  rev() |> 
  accumulate(~c(.x, .y))
xx
```
```{r message=FALSE}
degree <- lydian |> filter(role != "A4") |> pull("role") |> str_sub(2L)
res <- tibble(degree = degree,
              data = xx |> map(~lydian |> diminish(.x) |> as_12_TET()))
```

```{r}
res |> filter(degree == 5) |> pull("data")
```



## Use transformations to get 3rd, 5th and 7th, 9th, 11th and 14th

We can use indexing features to create a new index with period 7, and walk over it by steps of 2.

```{r}
# Ionian 1/12
demo <- start_with_mode() |>
  add_roles() |>
  as_12_TET()

# New index 2/7
demo |>  
  mutate(idx7 = index(dplyr::row_number() - 1, period = 7L, unit = 1L)) |> 
  xtibble(idx_name = "idx7", lock = FALSE) |>
  set_unit(to = 2L) |> 
  arrange(idx7)                                               
```

This works with any scale of 7 notes. Let's to this for the 7th degree of the melodic minor scale:

```{r}
demo <- get_data("melodic_minor") |>
  set_lock(to = FALSE) |>
  set_zero(on = 7, from = "degree") |>
  add_roles() |>
  as_12_TET()

demo |>
  mutate(idx7 = index(row_number() - 1, period = 7L, unit = 1L)) |> 
  xtibble(idx_name = "idx7", lock = FALSE) |> 
  set_unit(to = 2L) |> 
  arrange(idx7)                                               
```


And this can be scripted for an exhaustive demonstration:

```{r}
funky <- function(x, degree){
  x |>
    set_zero(on = degree, from = "degree") |>
    add_roles() |>
    as_12_TET() |>
    mutate(idx7 = index(row_number() - 1, period = 7L, unit = 1L)) |> 
    xtibble(idx_name = "idx7", lock = FALSE) |> 
    set_unit(to = 2L) |>
    select("idx7", "role") |>
    arrange(idx7)  
}
```

```{r}
mm <- get_data("melodic_minor") |> set_lock(to = FALSE, silently = TRUE)

demo <- mm |>
  select(degree) |>
  rowwise() |>
  mutate(data = funky(x = mm, degree = degree) |> list())
```
Unnest and pivot:

```{r}
demo |>
  unnest(data) |>
  mutate(idx7 = as.integer(idx7)) |>
  pivot_wider(names_from = idx7, values_from = role)
```
