---
title: "modes_5"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modes_5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(music)
library(dplyr)
library(tidyr)
library(purrr)
```


Some interersting reading:

- <https://en.wikipedia.org/wiki/Pentatonic_scale>


## Pentatonic as subsets of 7-notes modes

```{r}
demo <- match_modes()
demo
```

`delta` is just a shift on the circle of fifths.

A more insightful reasoning is to understand what the subset mean:

- major pentatonic is (P1-P5)-M2-M6-M3; it matches the 3 modes with a M3
- major blues is (P4-P1-P5)-M2-M6
- suspended is m7-(P4-P1-P5)-M2
- minor pentatonic is m3-m7-(P4-P1-P5): it matches the 3 minor modes with P5
- minor blues is m6-m3-m7-(P4-P1): it matches the 3 modes with a m6


Then it's a matter of presentation:

```{r}
demo |>
  select(delta, name_5, degree_7, name_7) |>
  pivot_wider(names_from = delta, values_from = name_5)
```
```{r}
demo |>
  select(delta, name_5, name_7) |>
  pivot_wider(names_from = delta, values_from = name_7)
```


## Particular pentonic

### Guitar tuning

E-A-D-G-D-E

This is G-D-A-E-B:

- G major
- D blues major
- A sus
- E minor
- B blues minor

### Black keys (piano)

F#/Gb major pentatonic.
