---
title: "modes_5"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modes_5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(music)
library(dplyr)
library(tidyr)
library(purrr)
```


More background:

- [Pentatonic scale](https://en.wikipedia.org/wiki/Pentatonic_scale)

Interesting things:

- No leading-tone, no subtonic (and no tritonic):
  - 2 is either 2M or 3m
  - 3 is either 3M or P4
  - 4 is either P5 or m6
  - 5 is either M6 or m7


- F maj, C maj, and G maj are the 3 white-keys transpositions


## TODO

Show the five sets of roles.

Show the natural triads which are only 3M and P4 stacks:

- 4-5 (degree 1 major): P1-M3-M6 : equivalently a minor triad inversion 3-4
- 5-4 (degree 5 blues major): P1-P4-M6 : equivalently a major triad inversion 4-3
- 5-5 (degrees 2, 3, 6): P1-P4-m7 : suspended 

Maybe with the circle (p5, u2) which has intervals 4-5-5-5(-5)...


## Pentatonic as subsets of 7-notes modes

```{r}
demo <- match_modes()
demo
```

`delta` is just the shift on the circle of fifths.

A more insightful reasoning is to understand what the subset mean:

- major pentatonic is (P1-P5)-M2-M6-M3:
    - no 4th and 7th
    - it matches the 3 modes with a M3
- major blues is (P4-P1-P5)-M2-M6:
    - no 3rd and 7th
- suspended is m7-(P4-P1-P5)-M2:
    - no 3rd and 6th
- minor pentatonic is m3-m7-(P4-P1-P5):
    - no 2nd and 6th
    - it matches the 3 minor modes with P5
- minor blues is m6-m3-m7-(P4-P1): 
  - no 2nd and 5th
  - it matches the 3 modes with a m6


Then it's a matter of presentation:

```{r}
demo |>
  select(delta, name_5, degree_7, name_7) |>
  pivot_wider(names_from = delta, values_from = name_5)
```
```{r}
demo |>
  select(delta, name_5, name_7) |>
  pivot_wider(names_from = delta, values_from = name_7)
```


## Particular pentonic scales

### Guitar tuning

E-A-D-G-D-E

This is G-D-A-E-B:

- G major
- D blues major
- A sus
- E minor
- B blues minor

### Black keys (piano)

- F#/Gb major pentatonic
- Eb minor pentatonic
