---
title: "music"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{music}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(music)
library(dplyr)
library(tidyr)
library(purrr)
```


This demonstrates typical pipes allowed by the package:

- start with a block of data;
- add complementary column;
- transform to a more commonly used coordinate system ("12_TET" = 12 tones equal temperament).

Additional steps can be inserted or appended.


## Start with some roles

```{r}
start_with_some_roles(c("m3", "P5", "m7")) |>
  add_notes(root_note = "D") |>
  as_12_TET()
```

## Start with some notes

```{r}
start_with_some_notes(c("C", "F", "A"), root_note = "D") |>
  add_roles() |>
  as_12_TET()
```

So this is a rootless Dm7.


### Inversion of a chord

After transforming, we ask to start on the second note.

```{r}
start_with_some_notes(c("C", "F", "A"), root_note = "D") |>
  add_roles() |>
  as_12_TET() |>
  set_first(rank = 2) |>
  mutate(interval = c(0, diff(idx)))
```
Note that this is still the same chord (with the same role).

Someone who's unfamiliar with chords and intervals might wonder what kind of "A" chord that could also be. The same 2 lines pipe as above could be run. Here we demonstrate how to go on from the previous result, after already made transformations. This requires to:

- unlock the `xtibble` object to use `set_zero()` and `set_unit_*()` on our index;
- transform back to an index ordered as fifths;
- set zero on "P5" from the "role" column (we can also equivalently set zero on "A" from "note");
- save the roles in another column (before adding new roles);
- add roles based on that new zero reference.

```{r}
start_with_some_notes(c("C", "F", "A"), root_note = "D") |>
  add_roles() |>
  as_12_TET() |>
  set_first(rank = 2) |>
  # new steps
  set_lock(to = FALSE, silently = TRUE) |>
  set_unit_to_fifths() |>
  set_zero("P5", from = "role_x") |> 
  rename("role_x" = "role") |>
  add_roles() |>
  as_12_TET()
```


### Advanced

We can explore similar questions by leveraging more programming skills. For example with `purrr` mapping functions and `tidyr` pivoting.

What are C, F and A when each is successively considered the root?

```{r}
demo <- c("C", "F", "A") |>
  {function(x){
    map_df(x, ~tibble(
      root = .x,
      data = start_with_some_notes(x, root_note = .x) |> add_roles() |> list()
    ))
  }}()

demo |> 
  unnest(data) |> 
  select(-idx) |>
  pivot_wider(names_from = root, names_prefix = "root:", values_from = role)
```

## Start with mode

See dedicated articles for 7 and 5 modes for more usage.


### Heptatonic scales:

```{r}
show_mode(mode_size = 7)
```


Dorian mode:

```{r}
start_with_mode(mode_size = 7, degree = 2) |>
  add_roles() |>
  as_12_TET()
```

### Pentatonic scales:

```{r}
show_mode(mode_size = 5)
```


## Alter roles

### Find other minor scales

There are several [minor scales](https://en.wikipedia.org/wiki/Minor_scale).

The harmonic minor scale is natural minor (aeolian) with M7 (raised m7):

```{r}
start_with_mode(degree = 6) |>
  add_roles() |>
  augment("m7")
```

The ascending melodic minor is:

- natural minor with M7 and M6 (harmonic minor with M6):

```{r}
start_with_mode(degree = 6) |>
  add_roles() |>
  augment(c("m7", "m6"))
```

- or equivalently dorian mode with M7:

```{r}
start_with_mode(degree = 2) |>
  add_roles() |>
  augment("m7")
```


### Blue notes:

Pentatonic minor with _d5_. Here we show how to add the role:

```{r}
start_with_mode(mode_size = 5, degree = 6) |>
  add_roles() |>
  pull("role") |>
  {\(x) c(x, "d5")}() |>
  start_with_some_roles() |>
  add_notes(root_note = "C") |>
  as_12_TET()
```

Blue notes on major scales: _m3_ and _m7_...

Some examples could be:

- major pentatonic with _m3_:

```{r}
start_with_mode(mode_size = 5, degree = 1) |>
  add_roles() |>
  pull("role") |>
  {\(x) c(x, "m3")}() |>
  start_with_some_roles() |>
  add_notes(root_note = "C") |>
  as_12_TET()
```

- ionian scale with alteration:
  - diminished 7th: myxolidian (m7) 
  - diminished 7th and 3d: dorian (m3, m7)
  - diminished 3d: ascending melodic minor (m3)

```{r}
start_with_mode(mode_size = 7, degree = 1) |>
  add_roles() |>
  diminish("M7") |>
  add_notes(root_note = "C") |>
  as_12_TET()
```


