---
title: "music"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{music}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(music)
library(dplyr)
library(tidyr)
library(purrr)
```


## Start with some roles

```{r}
start_with_some_roles(c("m3", "P5", "m7")) |>
  add_notes(root_note = "D") |>
  as_12_TET()
```

## Start with some notes

### Inversion of a chord

```{r}
start_with_some_notes(c("C", "F", "A"), root_note = "D") |>
  add_roles() |>
  as_12_TET() |>
  set_first(rank = 2) |>
  dplyr::mutate(interval = c(0, diff(idx)))
```
Compare:

```{r}
start_with_some_notes(c("C", "F", "A"), root_note = "D") |>
  add_roles() |>
  as_12_TET() |>
  set_first(rank = 2) |>
  dplyr::rename("role_x" = "role") |>
  set_lock(to = FALSE) |>
  set_zero("P5", from = "role_x") |> 
  set_unit_to_fifths() |>
  add_roles() |>
  as_12_TET()
```


### Advanced

What are F, A and C when each is successively considered the root?

```{r}
demo <- c("F", "A", "C") |>
  {function(x){
    purrr::map_df(x, ~tibble::tibble(
      root = .x,
      data = start_with_some_notes(x, root_note = .x) |> add_roles() |> list()
    ))
  }}()

demo |> 
  tidyr::unnest(data) |> 
  dplyr::select(-idx) |>
  tidyr::pivot_wider(names_from = root, names_prefix = "root:", values_from = role)
```

## Start with mode

Heptatonic scales:

```{r}
show_mode(mode_size = 7)
```

Pentatonic scales:

```{r}
show_mode(mode_size = 5)
```

Dorian:

```{r}
start_with_mode(mode_size = 7, degree = 2) |>
  add_roles() |>
  as_12_TET()
```


## Alter roles

### Blue notes:

Pentatonic minor with _d5_. Here we show how to add the role:

```{r}
start_with_mode(mode_size = 5, degree = 6) |>
  add_roles() |>
  pull("role") |>
  {\(x) c(x, "d5")}() |>
  start_with_some_roles() |>
  add_notes(root_note = "C") |>
  as_12_TET()
```

Blue notes on major scales: _m3_ and _m7_...

Some examples could be:

- major pentatonic with _m3_:

```{r}
start_with_mode(mode_size = 5, degree = 1) |>
  add_roles() |>
  pull("role") |>
  {\(x) c(x, "m3")}() |>
  start_with_some_roles() |>
  add_notes(root_note = "C") |>
  as_12_TET()
```

- ionian scale with alteration:
  - diminished 7th: myxolidian (m7) 
  - diminished 7th and 3d: dorian (m3, m7)
  - diminished 3d: ascending melodic minor (m3)

```{r}
start_with_mode(mode_size = 7, degree = 1) |>
  add_roles() |>
  diminish("M7") |>
  add_notes(root_note = "C") |>
  as_12_TET()
```

### Find other minor scales

There several [minor scales](https://en.wikipedia.org/wiki/Minor_scale).

The harmonic minor scale is natural minor (aeolian) with M7 (raised m7):

```{r}
start_with_mode(degree = 6) |>
  add_roles() |>
  augment("m7")
```

The ascending melodic minor is:

- natural minor with M7 and M6 (harmonic minor with M6)
- or dorian with M7

```{r}
start_with_mode(degree = 6) |>
  add_roles() |>
  augment(c("m7", "m6")) |>
  print()

start_with_mode(degree = 2) |>
  add_roles() |>
  augment("m7")
```

